---
title: "Preprocessing of raw data"
author: "Patrick Schratz"
output:
  html_notebook:
    code_folding: none
    fig_width: 10
    highlight: haddock
    number_sections: yes
    toc: yes
---

# Load packages

```{r, results='hide'}
library(tidyverse)
library(rio)
library(sf)
library(forcats)
library(stringr)
library(magrittr)
library(raster)
library(rgdal)
library(RSAGA)
library(GSIF)
library(lwgeom)
library(here)

source("99-funs.R")
```

# Load raw disease data

We have to 'wiggle' the x/y coordinates by a negligible amount so that they are unique; otherwise some models might will fail because the observations at identical locations must be identical.

```{r}
import("/data/patrick/raw/survey_data/diseases240112_mod.csv") %>% 
  as_tibble() -> disease_data

set.seed(1234)
disease_data$x <- disease_data$x + rnorm(nrow(disease_data))
disease_data$y <- disease_data$y + rnorm(nrow(disease_data))
```

# Export, re-locate and re-import

We moved 3 points a minimal amount in space because they are located right at the border of the lithology shapefile.
This was done in QGIS. 

Points that have been shifted:

- id = 718
- id = 649
- id = 1157

Subsequently, the following is the real "raw data".

```{r}
st_read("/data/patrick/mod/survey_data/points_mod.shp") %>% 
  as_tibble() -> disease_data
```

# Format date column to year only

Process `year` variable to end up with only years.

The two missing entries we added for observation `785` and `836` are based on statements from the person who collected the data.

```{r}
disease_data %<>% 
  mutate(year = fct_recode(date, "NA" = "-", 
                           "NA" = ""))

disease_data$year %<>% 
  str_replace_all("-","/") %>% 
  str_extract_all('\\d+') %>% 
  vapply(function(x) paste(x[3], collapse = '/'), character(1))

disease_data %<>% 
  mutate(year = replace(year, ID == 785, 2009)) %>% 
  mutate(year = replace(year, ID == 836, 2009))
```

# Change colnames to lowercase 

```{r}
disease_data %>% 
  colnames() %>% 
  tolower() -> colnames(disease_data)
```

# Filter to tree type "Pinus radiata"

We only analyze tree type "Pinus radiata" in this work.

```{r}
disease_data %<>% 
  filter(host == "Pinus radiata")
```

# Process variables

Here we transform the following environmental variables to be equal to [Iturritxa et al. (2014)](http://onlinelibrary.wiley.com/doi/10.1111/ppa.12328/abstract):

- `temp`: Mean temperature (March to September)
- `p_sum`: Precipitation sum (July - September)
- `r_sum`: Potential Incoming Solar Radiation (PISR) (July-August) as a relative fraction to its mean 

## Temperature (°C)

Reference: http://www.aemet.es/documentos/es/conocermas/recursos_en_linea/publicaciones_y_estudios/publicaciones/Atlas-climatologico/Atlas.pdf

Convert to Degree Celsius.

```{r}
disease_data %>% 
  dplyr::select(ttmarav, ttabrav, ttmayav, ttjunav, ttjulav, ttagoav, ttsepav) %>% 
  mutate_if(is.integer, as.numeric) %>% 
  rowMeans() %>% 
  divide_by(10) -> disease_data$temp
```

## Precipitation (mm/m^2)

Reference: http://www.aemet.es/documentos/es/conocermas/recursos_en_linea/publicaciones_y_estudios/publicaciones/Atlas-climatologico/Atlas.pdf

To remove outliers, we trim all values > 124.4 to 124.4.

```{r}
disease_data %>% 
  dplyr::select(pjuliav, pagosav, pseptav) %>% 
  rowSums() %>% 
  divide_by(10) -> disease_data$p_sum

disease_data %<>% 
  mutate(p_sum = replace(p_sum, p_sum < 124.4, 124.4))
```

## PISR 

Reference: http://www.aemet.es/documentos/es/conocermas/recursos_en_linea/publicaciones_y_estudios/publicaciones/Atlas-climatologico/Atlas.pdf

This variable is converted to the relative fraction of its mean (mean = 8003.512 kW/m^2).
Negative outliers are trimmed to -0.1.

```{r}
disease_data %>% 
  dplyr::select(rjulav, ragoav, rsepav) %>% 
  rowSums %>% 
  divide_by(mean(.)) - 1 -> disease_data$r_sum

disease_data %<>% 
  mutate(r_sum = replace(r_sum, r_sum < -0.1, -0.1)) 
```

## Slope

We use a Digitial Elevation Model (DEM) to extract the slope (percent) using the `RSAGA` package.
The resulting `.sgrd` file is converted into `.tif`.

Object `disease_data` is converted from a `data.frame` into a spatial object (`SpatialPointsDataFrame` (SPDF)) and projected to EPSG 32630.
The calculated slope values are read in as a raster object and extracted to the SPDF.

At the point of this analysis, the `raster `package was not able to deal with objects of class `sf` so multiple class transformations between raster* objects and sf objects were needed.

This variable was also trimmed to a value of 70 for the upper end.

```{r}
env <- rsaga.env(modules = "/usr/lib/x86_64-linux-gnu/saga/")
rsaga.slope.asp.curv(in.dem = "/data/patrick/mod/DEM/dem_5m.tif",
                     out.slope = "/data/patrick/mod/DEM/slope/slope_5m",
                     unit.slope = "degrees",
                     env = env)

rsaga.geoprocessor(lib = "io_gdal", module = 2, env = env,
                   param = list(GRIDS = "/data/patrick/mod/DEM/slope/slope_5m.sgrd", 
                                FILE = "/data/patrick/mod/DEM/slope/slope_5m.tif"))

disease_data %>% 
  st_as_sf() %>% 
  st_set_crs(32630) %>% 
  as("Spatial") -> disease_data_spdf

raster("/data/patrick/mod/DEM/slope/slope_5m.tif") %>% 
  raster::extract(disease_data_spdf) -> disease_data_spdf$slope_degrees

disease_data_spdf %>% 
  st_as_sf() -> disease_data

disease_data %<>% 
  mutate(slope_degrees = replace(slope_degrees, slope_degrees > 70, 70))
```

## Hail probability

Read in raster file.

```{r}
hail_ras <- raster("/data/patrick/raw/hail/Prob_GAM_square_area.tif")
```

Extract values to object `disease_data`.

```{r}
hail_ras %>% 
  raster::extract(disease_data) -> disease_data$hail_prob
```

## pH

- Spatial resolution: 5 km

- Projection: EPSG:3035 -> Lambert azimuthal equal-area 

The following code is mainly borrowed from [this GIS.Stackexchange](https://gis.stackexchange.com/questions/132403/how-to-read-adf-files-into-r?rq=1) question.

```{r}
new("GDALReadOnlyDataset", "/data/patrick/raw/ph_europe/ph_cacl2") %>% 
  asSGDF_GROD() %>% 
  raster() -> ph
```

Read in the study area polygon and crop the raster image to it.

For speed reasons, we first project the `pH` raster to the CRS of `study_area` and coerce it to the CRS of `disease_data` after the `crop()` call.

```{r , results='hide'}
st_read("/data/patrick/raw/boundaries/basque-country/Study_area.shp") %>% 
  st_transform(3035) %>% 
  as("Spatial") -> study_area

crs(ph) <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

ph %<>% 
  crop(study_area) %>% 
  mask(study_area) %>% 
  projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
```

Finally, extract values to `disease_data`.

```{r}
raster::extract(ph, disease_data) -> disease_data$ph
```

## Elevation

```{r}
raster("/data/patrick/mod/DEM/dem_5m.tif") %>% 
  raster::extract(disease_data) -> disease_data$elevation
```

## Lithology

Read, reproject and select desired column of lithology ressource.

```{r}
st_read("/data/patrick/raw/lithology/CT_LITOLOGICO_25000_ETRS89.shp", quiet = TRUE) %>% 
  st_set_crs("+proj=utm +zone=30 +ellps=GRS80 +units=m +no_defs") %>% 
  st_transform(32630) %>% 
  dplyr::select(COD_LITOLO) -> litho
```

Make it a valid polygon.
See also `?st_is_valid()`.

```{r}
st_make_valid(litho) -> litho_valid
```

Check if any points of `disease_data` do not intersect with the polygon.

```{r}
disease_data %>%
  st_intersects(litho_valid) %>% 
  lengths()
```

Do the intersection.

```{r}
disease_data %<>%
  st_intersection(litho_valid)
```

Change lithology entries to descriptive names.

Aggregated classes are based on common lithology (sub-)classes.

```{r}
disease_data %<>% 
  dplyr::rename(lithology = COD_LITOLO) %>% 
  mutate(lithology = as.factor(lithology)) %>% 
  mutate(lithology = fct_recode(lithology, 
                                "surface deposits" = "01", 
                                "clastic sedimentary rock" = "02",
                                "clastic sedimentary rock" = "03", 
                                "biological sedimentary rock" = "04", 
                                "clastic sedimentary rock" = "08", 
                                "biological sedimentary rock" = "09",
                                "biological sedimentary rock" = "10", 
                                "chemical sedimentary rock" = "11",
                                "chemical sedimentary rock" = "12", 
                                "magmatic rock" = "13",
                                "magmatic rock" = "14", 
                                "chemical sedimentary rock" = "15", 
                                "biological sedimentary rock" = "16", 
                                "biological sedimentary rock" = "17",
                                "clastic sedimentary rock" = "19", 
                                "magmatic rock" = "22",
                                "magmatic rock" = "23"
  ))
```

## Soil

Source: 250m SoilGrids (worldwide) by T. Hengl -> http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169748

Legend: http://gsif.r-forge.r-project.org/soil.legends.html

```{r}
data("soil.legends")

soil.legends$TAXNWRB
```

Crop and reproject `soil` to CRS of soil ressource.
Convert to CRS of `disease_data` after `crop()` and `mask()` operations for speed reasons.

```{r}
st_read("/data/patrick/raw/boundaries/basque-country/Study_area.shp") %>% 
  st_transform("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") %>% 
  as("Spatial") -> study_area


raster("/data/patrick/raw/soil/ISRIC_world_soil_information/TAXNWRB_250m_ll.tif") %>% 
  crop(study_area) %>% 
  mask(study_area) %>% 
  projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs") -> soil
```

Some raster cells have no integer values but something like '27.49932' so we converted such values to Integers (`as.integer()`).

Two points have no raster coverage:

* Point 1: Located outside the raster coverage (northern coast).
* Point 2: Missing data in the centre of the study area .

We did a manual assignment based on nearest soil class ('27' in both cases).

```{r}
raster::extract(soil, disease_data) -> disease_data$soil

disease_data %<>% 
  mutate(soil = replace(soil, id == 245 | id == 442, 27)) 
```

Change soil classes to descriptive names.

```{r}
disease_data %<>% 
  st_as_sf() %>% 
  mutate(soil = as.factor(as.integer(soil))) %>% 
  mutate(soil = fct_recode(soil, 
                           "Cambisols" = "27", "Cambisols" = "26", 
                           "Cambisols" = "28", "Cambisols" = "29", 
                           "Cambisols" = "30", "Cambisols" = "31",
                           "Cambisols" = "32", "Cambisols" = "33", 
                           "Cambisols" = "34", "Cambisols" = "35", 
                           "Chernozems" = "36", "Chernozems" = "37",
                           "Chernozems" = "38", "Cryosols" = "39", 
                           "Cryosols" = "40", "Cryosols" = "41", 
                           "Durisols" = "42", "Ferralsols" = "43", 
                           "Ferralsols" = "44", "Ferralsols" = "45", 
                           "Ferralsols" = "46", "Ferralsols" = "47", 
                           "Fluvisols" = "49", "Fluvisols" = "51", 
                           "Fluvisols" = "52", "Gleysols" = "53", 
                           "Gleysols" = "54", "Gleysols" = "55", 
                           "Gleysols" = "56", "Gleysols" = "57", 
                           "Gleysols" = "58", "Gypsisols" = "59",
                           "Histosols" = "61", "Histosols" = "62",
                           "Histosols" = "63", "Histosols" = "64", 
                           "Histosols" = "65", "Kastanozems" = "66", 
                           "Kastanozems" = "67", "Leptosols" = "68", 
                           "Leptosols" = "69", "Leptosols" = "70", 
                           "Leptosols" = "71", "Leptosols" = "72", 
                           "Lixisols" = "73", "Lixisols" = "74", 
                           "Lixisols" = "75", "Luvisols" = "76", 
                           "Luvisols" = "77", "Luvisols" = "78", 
                           "Luvisols" = "79"
                           ))
```

Aggregate soil classes.

Reference: http://www.fao.org/3/a-i3794e.pdf
http://www.fao.org/3/i3794en/I3794en.pdf

```{r}
disease_data %<>% 
  st_as_sf() %>% 
  mutate(soil = fct_recode(soil, 
                           "soils with limitations to root growth" = "Leptosols",
                           "soils with limitations to root growth" = "Cryosols",
                           "soils with clay-enriched subsoil" = "Luvisols",
                           "soils with clay-enriched subsoil" = "Lixisols",
                           "organic soi" = "Histosols",
                           "pronounced accumulation of organic matter in the mineral topsoil" = "Kastanozems",
                           "pronounced accumulation of organic matter in the mineral topsoil" = "Chernozems",
                           "accumulation of moderately soluble salts or non-saline substances" = "Gypsisols",
                           "accumulation of moderately soluble salts or non-saline substances" = "Durisols",
                           "soils distinguished by Fe/Al chemistry" = "Gleysols",
                           "soils with little or no profile differentiation" = "Fluvisols",
                           "soils with little or no profile differentiation" = "Cambisols",
                           "soils distinguished by Fe/Al chemistry" = "Ferralsols"
                           ))
```

## Assign coordinates as variables

We need the coordinates as variables here for the age imputation.

```{r}
disease_data$x <- as.integer(st_coordinates(disease_data)[, 1])
disease_data$y <- as.integer(st_coordinates(disease_data)[, 2])
```

## Age 

We imputeted variable `age`.

```{r}
disease_data %<>% 
  filter(!is.na(age))
```

We trimmed age > 40 to reduce influence of outliers.

```{r}
disease_data %<>% 
  mutate(age = as.numeric(levels(age)[age])) %>% 
  mutate(age = replace(age, age > 40, 40))
```

```{r}
res <- test_age_imputation(disease_data$x, disease_data$y, disease_data$age)
summary(res)
```

```{r}
qplot(res$obs, res$imput) + geom_abline() + hrbrthemes::theme_ipsum_rc()
cor(res$obs, res$imput, use = "complete")
```

```{r}
disease_data$age <- age_imputation(disease_data$x, disease_data$y, 
                                   disease_data$age, print = TRUE)
```

```{r}
disease_data %<>% 
  filter(!is.na(age))
```

# Sort dataframe

```{r}
disease_data %<>% 
  dplyr::select(x, y, diplo01, temp, p_sum, r_sum, elevation, 
                slope_degrees, hail_prob, age, ph, lithology, soil, year)
```

# Transform variables and save

```{r}
disease_data %<>% 
  as.tibble() %>% 
  droplevels() %>% 
  mutate(diplo01 = as.factor(diplo01)) %>% 
  mutate(year = as.factor(year))

saveRDS(disease_data, "/data/patrick/mod/survey_data/data-clean.rda")

# write shapefile for study area plotting
# disease_data %>% 
#   st_as_sf(coords = c("x", "y"), crs = 32630) %>% 
#   st_write(here("05_data/data-clean.gpkg"), layer = "data-clean.gpkg") 

#file.copy(here("05_data/data-clean.gpkg"), "/data/patrick/test.gpkg")

#test <- st_read("/data/patrick/test.gpkg")
```

# Session Info

```{r}
sessionInfo()
```

