---
title: "Spatial prediction"
author: Patrick Schratz
output:
  html_notebook:
    code_folding: none
    fig_width: 10
    highlight: haddock
    number_sections: yes
    toc: yes
---

Using the GAM as the prediction maps will be much smoother than using RF and the predictive performance of both models was equal.

# Load packages

```{r}
library(devtools)
library(sf)
library(magrittr)
library(raster)
library(purrr)
library(rgdal)
library(dplyr)
library(lwgeom)
library(forcats)
library(tibble)
library(viridis)
library(mapview)
library(rasterVis)
library(lattice)
library(latticeExtra)

install_github("pat-s/mlr@all-mods")
library(mlr)
```

# Load data

```{r}
df <- readRDS("/data/patrick/mod/survey_data/data-clean.rda")
df$geometry <- NULL
df$age <- NULL # age is not spatially available
df$year = NULL
```

# Create Task

```{r}
task_spatial <- makeClassifTask(id = "pathogen_data", data = df,
                                target = "diplo01", positive = "1",
                                spatial = TRUE)
```

# Create learners

```{r}
lrn_gam <- makeLearner("classif.gam",
                       family = "binomial",
                       binomial.link = "logit",
                       predict.type = "prob",
                       method = "GCV.Cp",
                       fix.factors.prediction = TRUE,
                       formula = "diplo01 ~ s(temp, k = 10) + s(p_sum, k = 10) + s(r_sum, k = 10) +
                       s(elevation, k = 10) + s(slope_degrees, k = 10) + s(hail_prob, k = 10) + s(ph, k = 10) +
                       lithology + soil")

lrn_rf <- makeLearner("classif.ranger", 
                      predict.type = "prob",
                      mtry = 1)
```

# Train model

```{r}
if (!file.exists("/data/patrick/mod/modeling/paper1_2018/gam_prediction.rda")) {
  model <- train(lrn_gam, task_spatial)
  model_rf <- train(lrn_rf, task_spatial)
  saveRDS(model, "/data/patrick/mod/modeling/paper1_2018/gam_prediction.rda")
  saveRDS(model_rf, "/data/patrick/mod/modeling/paper1_2018/rf_prediction.rda")
} else {
  model = readRDS("/data/patrick/mod/modeling/paper1_2018/gam_prediction.rda")
  model_rf = readRDS("/data/patrick/mod/modeling/paper1_2018/rf_prediction.rda")
}
```

# Create prediction data

## Load study area

Its a square around the Basque Country. 
This prevents error for the subsequent raster `extract` calls that may result in NAs if the rasters does not exceed the dimensions of the extracting area.

```{r}
study_area <- st_read("/data/patrick/raw/boundaries/basque-country/Study_area.shp") %>% 
  st_transform(32630) %>% 
  as("Spatial")
```

## Temperature (Mar - Sep)

We load all images for the months March - September. 
These are clipped in a rectangular box around the Basque Country.

```{r}
# load only nov-april
rasters_temp = list.files("/data/patrick/master_thesis/AtlasClimatico/04_Tiffs_sub/",
                     pattern = "temp_mean_03|temp_mean_04|temp_mean_05|temp_mean_06|temp_mean_07|temp_mean_08|temp_mean_09", 
                     full.names = TRUE)
```

We load in each raster and set the projection.
Next, we mask the values by the study area extent and rescale the temperature values to have proper degree Celsius values.

Then a stack is created from all the rasters and the mean is computed across all layers.

```{r}
if (!file.exists("/data/patrick/mod/prediction/temporary/rasters_temp_mean.rda")) {
  # set up list with monthly tifs and rescale values 
  rasters_temp %<>% map(~ {
    tmp <- raster(.x, layer = 1)
    # apply CRS
    crs(tmp) <- "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"
    # subset and rescale 
    tmp %<>% 
      crop(study_area) %>% 
      mask(study_area) %>% 
      divide_by(10)
    return(tmp)
  })
  
  rasters_temp_stack = stack(rasters_temp)
  rasters_temp_mean <- stackApply(rasters_temp_stack, 
                                  indices = c(rep(1, 7)), fun = mean)
  
  saveRDS(rasters_temp_mean, "/data/patrick/mod/prediction/temporary/rasters_temp_mean.rda")
} else {
  rasters_temp_mean <- readRDS("/data/patrick/mod/prediction/temporary/rasters_temp_mean.rda")
}
```

## Prediction raster grid

Now create a 200 x 200 point grid using a dummy raster that represents the study area shape.

!! This is temporarly overwritten by an import of the below created file which has already the extracted ltihology information inisde. 
This intersection was created in ArcGIS as R did not finish in one day. 
Runtime in ArcGIS: < 1 min.
QGIS did not even start..

```{r}
df_basque <- st_read("/data/patrick/mod/prediction/temporary/basque_litho.shp")
df_basque$OBJECTID = NULL
df_basque$layer = NULL
df_basque$FID_df_bas = NULL
df_basque$slp_dgr = NULL
df_basque$FID_litho = NULL
```


```{r}
# df_basque <- as(rasters_temp_mean, 'SpatialPixelsDataFrame') %>% 
#   st_as_sf() 
# df_basque$index_1 <- NULL
```

And extract the information to the new data set.

```{r}
df_basque$temp <- raster::extract(rasters_temp_mean, df_basque)
```

## Precipitation (Mar - Sep)

We load all images for the months July - September. 
These are clipped in a rectangular box around the Basque Country.

```{r}
# load only nov-april
rasters_prec = list.files("/data/patrick/master_thesis/AtlasClimatico/04_Tiffs_sub/",
                     pattern = "prec_r_07|prec_r_08|prec_r_09", 
                     full.names = TRUE)
```

We load in each raster and set the projection.
Next, we mask the values by the study area extent and rescale the precerature values to have mm/m² values.

Then a stack is created from all the rasters and the mean is computed across all layers.

```{r}
if (!file.exists("/data/patrick/mod/prediction/temporary/rasters_p_sum.rda")) {
  # set up list with monthly tifs and rescale values 
  rasters_prec %<>% map(~ {
    tmp <- raster(.x, layer = 1)
    # apply CRS
    crs(tmp) <- "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"
    # subset and rescale 
    tmp %<>% 
      crop(study_area) %>% 
      mask(study_area) %>% 
      divide_by(10)
    return(tmp)
  })
  
  rasters_prec_stack = stack(rasters_prec)
  rasters_prec_sum <- stackApply(rasters_prec_stack, 
                                  indices = c(rep(1, 3)), fun = sum)
  
  saveRDS(rasters_prec_sum, "/data/patrick/mod/prediction/temporary/rasters_p_sum.rda")
} else {
  rasters_prec_sum <- readRDS("/data/patrick/mod/prediction/temporary/rasters_p_sum.rda")
}
```

```{r}
df_basque$p_sum <- raster::extract(rasters_prec_sum, df_basque)
```

## PISR (Mar - Sep)

We load all images for the months July - September. 
These are clipped in a rectangular box around the Basque Country.

```{r}
rasters_r_sum = list.files("/data/patrick/master_thesis/AtlasClimatico/04_Tiffs_sub/",
                           pattern = "sRad_07|sRad_08|sRad_09", 
                           full.names = TRUE)
```

We load in each raster and set the projection.
Next, we mask the values by the study area extent and rescale the sRaderature values to have mm/m² values.

Then a stack is created from all the rasters and the mean is computed across all layers.

```{r}
if (!file.exists("/data/patrick/mod/prediction/temporary/rasters_r_sum.rda")) {
  # set up list with monthly tifs and rescale values 
  rasters_r_sum %<>% map(~ {
    tmp <- raster(.x, layer = 1)
    # apply CRS
    crs(tmp) <- "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"
    # subset and rescale 
    tmp %<>% 
      crop(study_area) %>% 
      mask(study_area)
      #divide_by(mean(.)) - 1
    return(tmp)
  })
  
  rasters_r_sum_stack = stack(rasters_r_sum)
  rasters_r_sum <- stackApply(rasters_r_sum_stack, 
                              indices = c(rep(1, 3)), fun = sum)
  
  # trim outliers; we have around 50000 values with no information that affect the mean
  # these are coming from the sea area where no information is available
  rasters_r_sum <- calc(rasters_r_sum, fun = function(x) {replace(x, x < 2000, NA)})
  
  # rescale to fraction its mean now and remove NA values when calc the mean
  rasters_r_sum <- calc(rasters_r_sum, fun = function(x) {x / mean(x, na.rm = TRUE) - 1})
  
  # again trim negative outliers
  rasters_r_sum <- calc(rasters_r_sum, fun = function(x) {replace(x, x < -0.1, -0.1)})
  
  saveRDS(rasters_r_sum, "/data/patrick/mod/prediction/temporary/rasters_r_sum.rda")
} else {
  rasters_r_sum <- readRDS("/data/patrick/mod/prediction/temporary/rasters_r_sum.rda")
}
```

```{r}
df_basque$r_sum <- raster::extract(rasters_r_sum, df_basque)
```

## Slope

We use a Digitial Elevation Model (DEM) to extract the slope (percent) using the `RSAGA` package.
The resulting `.sgrd` file is converted into `.tif`.

Object `disease_data` is converted from a `data.frame` into a spatial object (`SpatialPointsDataFrame` (SPDF)) and projected to EPSG 32630.
The calculated slope values are read in as a raster object and extracted to the SPDF.

At the point of this analysis, the `raster `package was not able to deal with objects of class `sf` so multiple class transformations between raster* objects and sf objects were needed.

This variable was also trimmed to a value of 70 for the upper end.

First we create a 200x200 point grid using one of the already existing raster files.

```{r}
if (!file.exists("/data/patrick/mod/DEM/slope/slope_5m.tif")) {
  
  env <- rsaga.env(modules = "/usr/lib/x86_64-linux-gnu/saga/")
  rsaga.slope.asp.curv(in.dem = "/data/patrick/mod/DEM/dem_5m.tif",
                       out.slope = "/data/patrick/mod/DEM/slope/slope_5m",
                       unit.slope = "degrees",
                       env = env)
  
  rsaga.geoprocessor(lib = "io_gdal", module = 2, env = env,
                     param = list(GRIDS = "/data/patrick/mod/DEM/slope/slope_5m.sgrd", 
                                  FILE = "/data/patrick/mod/DEM/slope/slope_5m.tif"))
} else {
  dem <- raster("/data/patrick/mod/DEM/slope/slope_5m.tif")
}
dem %>% 
  raster::extract(df_basque) -> df_basque$slope_degrees
```

## Lithology

Read, reproject and select desired column of lithology ressource.

```{r}
# st_read("/data/patrick/raw/lithology/CT_LITOLOGICO_25000_ETRS89.shp", quiet = TRUE) %>%
#   st_set_crs("+proj=utm +zone=30 +ellps=GRS80 +units=m +no_defs") %>%
#   st_transform(32630) %>%
#   select(COD_LITOLO) -> litho
```

Make it a valid polygon.
See also `?st_is_valid()`.

```{r}
#st_make_valid(litho) -> litho_valid
```

Do the intersection.

```{r}
# df_basque %<>%
#   st_intersection(litho_valid)
```

Change lithology entries to descriptive names.

Aggregated classes are based on common lithology (sub-)classes.

```{r}
df_basque %<>% 
  dplyr::rename(lithology = COD_LITOLO) %>% 
  mutate(lithology = as.factor(lithology)) %>% 
  mutate(lithology = fct_recode(lithology, 
                                NULL = "00", # water bodies
                                "surface deposits" = "01", 
                                "clastic sedimentary rock" = "02",
                                "clastic sedimentary rock" = "03", 
                                "biological sedimentary rock" = "04", 
                                "clastic sedimentary rock" = "08", 
                                "biological sedimentary rock" = "09",
                                "biological sedimentary rock" = "10", 
                                "chemical sedimentary rock" = "11",
                                "chemical sedimentary rock" = "12", 
                                "magmatic rock" = "13",
                                "magmatic rock" = "14", 
                                "chemical sedimentary rock" = "15", 
                                "biological sedimentary rock" = "16", 
                                "biological sedimentary rock" = "17",
                                "dolomites" = "18",
                                "clastic sedimentary rock" = "19", 
                                "magmatic rock" = "20",
                                "magmatic rock" = "22",
                                "magmatic rock" = "23",
                                NULL = "24" # riffs
  ))
```

## Soil

```{r}
st_read("/data/patrick/raw/boundaries/basque-country/Study_area.shp") %>% 
  st_transform("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") %>% 
  as("Spatial") -> study_area


raster("/data/patrick/raw/soil/ISRIC_world_soil_information/TAXNWRB_250m_ll.tif") %>% 
  crop(study_area) %>% 
  mask(study_area) %>% 
  projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs") -> soil
```

```{r}
df_basque$soil <- raster::extract(soil, df_basque)
```

```{r}
df_basque %<>% 
  st_as_sf() %>% 
  mutate(soil = as.factor(as.integer(soil))) %>% 
  mutate(soil = fct_recode(soil, 
                           "Acrisols" = "2", "Acrisols" = "2",
                           "Albeluvisol" = "7",
                           "Andosols" = "14", "Arenosols" = "16",
                           "Arenosols" = "18", "Protic" = "20",
                           "Calcisols" = "21", "Calcisols" = "22",
                           "Calcisols" = "23", "Calcisols" = "24",
                           "Cambisols" = "25",
                           "Cambisols" = "27", "Cambisols" = "26", 
                           "Cambisols" = "28", "Cambisols" = "29", 
                           "Cambisols" = "30", "Cambisols" = "31",
                           "Cambisols" = "32", "Cambisols" = "33", 
                           "Cambisols" = "34", "Cambisols" = "35", 
                           "Chernozems" = "36", "Chernozems" = "37",
                           "Chernozems" = "38", "Cryosols" = "39", 
                           "Cryosols" = "40", "Cryosols" = "41", 
                           "Durisols" = "42", "Ferralsols" = "43", 
                           "Ferralsols" = "44", "Ferralsols" = "45", 
                           "Ferralsols" = "46", "Ferralsols" = "47", 
                           "Fluvisols" = "48",
                           "Fluvisols" = "49", "Fluvisols" = "50",
                           "Fluvisols" = "51", 
                           "Fluvisols" = "52", "Gleysols" = "53", 
                           "Gleysols" = "54", "Gleysols" = "55", 
                           "Gleysols" = "56", "Gleysols" = "57", 
                           "Gleysols" = "58", "Gypsisols" = "59",
                           "Gypsisols" = "60", 
                           "Histosols" = "61", "Histosols" = "62",
                           "Histosols" = "63", "Histosols" = "64", 
                           "Histosols" = "65", "Kastanozems" = "66", 
                           "Kastanozems" = "67", "Leptosols" = "68", 
                           "Leptosols" = "69", "Leptosols" = "70", 
                           "Leptosols" = "71", "Leptosols" = "72", 
                           "Lixisols" = "73", "Lixisols" = "74", 
                           "Lixisols" = "75", "Luvisols" = "76", 
                           "Luvisols" = "77", "Luvisols" = "78", 
                           "Luvisols" = "79", "Luvisols" = "83",
                           "Luvisols" = "84", "Nitisols" = "86",
                           "Phaeozems" = "87", "Phaeozems" = "88",
                           "Planosols" = "90", "Planosols" = "91",
                           "Planosols" = "93", "Planosols" = "94",
                           "Plinthosols" = "95", "Podzols" = "97", 
                           "Podzols" = "98", "Vertisols" = "116"
                           ))
```

```{r}
df_basque %<>% 
  st_as_sf() %>% 
  mutate(soil = fct_recode(soil, 
                           "soils with limitations to root growth" = "Leptosols",
                           "soils with limitations to root growth" = "Cryosols",
                           "soils with limitations to root growth" = "Vertisols",
                           "soils with clay-enriched subsoil" = "Luvisols",
                           "soils with clay-enriched subsoil" = "Lixisols",
                           "soils with clay-enriched subsoil" = "Acrisols",
                           "organic soil" = "Histosols",
                           "pronounced accumulation of organic matter in the mineral topsoil" = "Kastanozems",
                           "pronounced accumulation of organic matter in the mineral topsoil" = "Chernozems",
                           "pronounced accumulation of organic matter in the mineral topsoil" = "Phaeozems",
                           "accumulation of moderately soluble salts or non-saline substances" = "Gypsisols",
                           "accumulation of moderately soluble salts or non-saline substances" = "Durisols",
                           "accumulation of moderately soluble salts or non-saline substances" = "Calcisols",
                           "soils distinguished by Fe/Al chemistry" = "Gleysols",
                           "soils distinguished by Fe/Al chemistry" = "Podzols",
                           "soils distinguished by Fe/Al chemistry" = "Plinthosols",
                           "soils distinguished by Fe/Al chemistry" = "Planosols",
                           "soils distinguished by Fe/Al chemistry" = "Nitisols",
                           "soils with little or no profile differentiation" = "Fluvisols",
                           "soils with little or no profile differentiation" = "Cambisols",
                           "soils with little or no profile differentiation" = "Arenosols",
                           "soils distinguished by Fe/Al chemistry" = "Ferralsols"
                           ))
```

## Hail probability

Read in raster file.

```{r}
hail_ras <- raster("/data/patrick/raw/hail/Prob_GAM_square_area.tif")
```

Extract values to object `df_basque`.

```{r}
hail_ras %>% 
  raster::extract(df_basque) -> df_basque$hail_prob
```

## Elevation

```{r}
raster("/data/patrick/mod/DEM/dem_5m.tif") %>% 
  raster::extract(df_basque) -> df_basque$elevation
```

## pH

```{r}
new("GDALReadOnlyDataset", "/data/patrick/raw/ph_europe/ph_cacl2") %>% 
  asSGDF_GROD() %>% 
  raster() -> ph
```

Read in the study area polygon and crop the raster image to it.

For speed reasons, we first project the `pH` raster to the CRS of `study_area` and coerce it to the CRS of `df_basque` after the `crop()` call.

```{r , results='hide'}
st_read("/data/patrick/raw/boundaries/basque-country/Study_area.shp") %>% 
  st_transform(3035) %>% 
  as("Spatial") -> study_area

crs(ph) <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

ph %<>% 
  crop(study_area) %>% 
  mask(study_area) %>% 
  projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
```

Finally, extract values to `disease_data`.

```{r}
raster::extract(ph, df_basque) -> df_basque$ph
```

## Remove NAs

```{r}
df_basque %<>% na.omit()
```

# Predict

Make prediction data.frame

```{r}
df_basque_pred <- df_basque
df_basque_pred$geometry = NULL
```

## GAM

```{r}
pred <- predict(model, newdata = df_basque_pred)
```

## RF

```{r}
pred_rf <- predict(model_rf, newdata = df_basque_pred)
```

# Assign coordinates and visualize

## GAM

```{r}
coords <- tibble(x = as.integer(st_coordinates(df_basque)[, 1]),
                 y = as.integer(st_coordinates(df_basque)[, 2]))
# not sure why we have some non-matching row names here ??
pred_xy = left_join(rownames_to_column(pred$data), 
                    rownames_to_column(coords), by = "rowname") %>% 
  na.omit()
pred_xy = SpatialPixelsDataFrame(points = pred_xy[, c("x", "y")], 
                                 data = pred_xy, 
                                 proj4string = CRS("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"))

mapview(pred_xy, legend = TRUE, zcol = "prob.1", 
                 at = c(0, 0.25, 0.5, 0.75, 1), layer.name = "prob.1")

# write out raster
raster_gam <- stack(pred_xy)
raster_gam <- projectRaster(raster_gam, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")

writeRaster(raster_gam, "/data/patrick/mod/prediction/diplodia_gam.tif",
            overwrite = TRUE)
```

## RF

```{r}
coords <- tibble(x = as.integer(st_coordinates(df_basque)[, 1]),
                 y = as.integer(st_coordinates(df_basque)[, 2]))
# not sure why we have some non-matching row names here ??
pred_xy_rf = left_join(rownames_to_column(pred_rf$data), 
                       rownames_to_column(coords), by = "rowname") %>% 
  na.omit()
pred_xy_rf = SpatialPixelsDataFrame(points = pred_xy_rf[, c("x", "y")], 
                                    data = pred_xy_rf, 
                                    proj4string = CRS("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"))

mapview(pred_xy_rf, legend = TRUE, zcol = "prob.1", 
                 at = c(0, 0.25, 0.5, 0.75, 1), layer.name = "prob.1")

# write out raster
raster_rf <- stack(pred_xy_rf)
raster_rf <- projectRaster(raster_rf, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs ")

writeRaster(raster_rf, "/data/patrick/mod/prediction/diplodia_rf.tif",
            overwrite = TRUE)
```

# Create map

## GAM

```{r}
study_area %<>% 
  st_as_sf() %>% 
  st_transform(4326) %>% 
  as("Spatial")

map_gam = levelplot( 
  raster_gam[["prob.1"]], 
  main = substitute(paste("Probability of tree infections by ", 
                          italic('Diplodia pinea'))), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0),          
  colorkey = list(
    width = 0.75,                       
    space = 'right',                     
    labels = list(at = c(0, 0.25, 0.5, 0.75, 1),
                font = 4),      
    axis.line = list(col='black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis,                  
  at = c(0, 0.25, 0.5, 0.75, 1)) +             
  layer(sp.polygons(study_area, lwd = 10))                 
```

```{r}
png("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/prediction_gam.png")
map_gam
dev.off()
```

## RF

```{r}
study_area %<>% 
  st_as_sf() %>% 
  st_transform(4326) %>% 
  as("Spatial")

map_rf = levelplot( 
  raster_rf[["prob.1"]], 
  main = substitute(paste("Probability of tree infections by ", 
                          italic('Diplodia pinea'))), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0),          
  colorkey = list(
    width = 0.75,                       
    space = 'right',                     
    labels = list(at = c(0, 0.25, 0.5, 0.75, 1),
                font = 4),      
    axis.line = list(col = 'black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis,                  
  at = c(0, 0.25, 0.5, 0.75, 1)) +             
  layer(sp.polygons(study_area, lwd = 10))                 
```

```{r}
png("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/prediction_rf.png")
map_rf
dev.off()
```
