---
title: "R Notebook"
output: html_notebook
---

# Load packages and data

```{r}
pacman::p_load(car, tidyverse, vcd, ggplot2, hrbrthemes, reporttools)
source("99-funs.R")
```

```{r}
readRDS("/home/shares/data/LIFE/mod/survey_data/data-clean.rda") %>% 
  as_tibble() -> df
```

# check for correlation of variables

## continuous variables

### 'my_pairs()' OK

elevation + ph:   0.55 
temp + p_sum:     0.31
p_sum + slope:    0.24

```{r }
df %>% 
  dplyr::select(temp, p_sum, r_sum, elevation, slope, age, ph) %>% 
  my_pairs(cex.cor = 2)
```

### corVIF OK

```{r }
vif(glm(diplo01 ~ temp + p_sum + r_sum + elevation + slope + age + ph, data = df,
             family = binomial(link = 'logit')))
```

## categorical variables OH?

### Cramer's V

- soil + lithology: 0.15
- soil + hail:      0.117
- lithology + hail: 0.578 (oh?) -> but does a correlation here make any sense? Probably fake-correlation

https://stats.stackexchange.com/a/236056/101464

```{r}
vars <- c("soil", "lithology", "hail", "hail_new")
catcorrm <- function(vars, dat) sapply(vars, function(y) 
  sapply(vars, function(x) assocstats(table(dat[,x], dat[,y]))$cramer)) 

catcorrm(vars, as.data.frame(df))
```

## continuous and categorical

### Intraclass correlcation coefficient (ICC)

```{r}
df %>% 
  dplyr::select(temp, p_sum, r_sum, elevation, slope, age, ph) %>%
  ICC()
```

# Some plots

## temp vs p_sum

```{r}
ggplot(data = df, aes(x = temp, y = p_sum)) + 
  geom_point() + 
  theme_ipsum_rc()
```

## temp vs r_sum
```{r}
ggplot(data = df, aes(x = temp, y = r_sum)) + 
  geom_point() + 
  theme_ipsum_rc()
```
## elevation vs p_sum
```{r}
ggplot(data = df, aes(x = elevation, y = p_sum)) + 
  geom_point() + 
  theme_ipsum_rc()
```

## elevation vs r_sum
```{r}
ggplot(data = df, aes(x = elevation, y = r_sum)) + 
  geom_point() + 
  theme_ipsum_rc()
```

## hail vs temp
```{r}
ggplot(data = df, aes(x = hail, y = temp)) + 
  geom_boxplot() + 
  theme_ipsum_rc()
```

## hail vs p_sum
```{r}
ggplot(data = df, aes(x = hail, y = p_sum)) + 
  geom_boxplot() + 
  theme_ipsum_rc()
```

## ph vs soil 

```{r}
ggplot(data = df, aes(x = soil, y = ph)) + 
  geom_boxplot() + 
  theme_ipsum_rc()
```

## ph vs lithology

```{r}
ggplot(data = df, aes(x = lithology, y = ph)) + 
  geom_boxplot() + 
  theme_ipsum_rc()
```

## diplo vs hail

```{r}
ggplot(df, aes(hail, ..count..)) + 
  geom_bar(aes(fill = diplo01), position = "dodge") +
  theme_ipsum_rc()
```

## fus vs hail

```{r}
ggplot(df, aes(hail, ..count..)) + 
  geom_bar(aes(fill = fus01), position = "dodge") +
  theme_ipsum_rc()
```

## diplo vs soil

```{r}
ggplot(df, aes(soil, ..count..)) + 
  geom_bar(aes(fill = diplo01), position = "dodge") +
  theme_ipsum_rc()
```

## fus vs soil

```{r}
ggplot(df, aes(soil, ..count..)) + 
  geom_bar(aes(fill = fus01), position = "dodge") +
  theme_ipsum_rc()
```

## diplo vs litho

```{r}
ggplot(df, aes(lithology, ..count..)) + 
  geom_bar(aes(fill = diplo01), position = "dodge") +
  theme_ipsum_rc()
```

## fus vs litho

```{r}
ggplot(df, aes(lithology, ..count..)) + 
  geom_bar(aes(fill = fus01), position = "dodge") +
  theme_ipsum_rc()
```

# Descriptive tables of data

```{r}
readRDS("/home/shares/data/LIFE/mod/survey_data/data-clean.rda") %>% 
  as_tibble() %>% 
  as.data.frame() -> df

df$geometry <- NULL
df$fus01 <- NULL
df$id <- NULL
df$y <- NULL
df$x <- NULL
df$hail <- NULL

# summarize numeric variables
tableContinuous(df[, sapply(df, is.numeric)],
                longtable = FALSE, cumsum = FALSE,
                cap = "Descriptive summary of numerical variables", 
                label = "table:descriptive_summary_numeric")
```

```{r}
# summarize non-numeric variables
tableNominal(df[, !sapply(df, is.numeric)],
             longtable = FALSE, cumsum = FALSE,
             cap = "Descriptive summary of non-numerical variables",
             label = "table:descriptive_summary_non_numeric")
```

