---
title: "Evaluating classifiers on pathogens"
output: html_notebook
author: "Patrick Schratz, FSU Jena"
---

This report evaluates the performance of the algorithms 

* _Boosted Regression Trees_ (BRT), 
* _Random Forest_ (RF), 
* _Support Vector Machine_ (SVM) 
* _Extreme Gradient Boosting_ (xgboost) 
* _Generalized Linear Model_ (GLM) 
* _Generalized Additive Model_ (GAM) 

on four different pathogens:

- _Armillaria mellea_
- _Diplodia sapinea_
- _Fusarium mellea_
- _Heterobasidion mellea_

```{r setup, include=FALSE}
drake::loadd(bm_sp_sp_all)
# knitr::opts_chunk$set(eval = FALSE)
```

# Resampling strategies {.tabset .tabset-fade}

## Spatial-Spatial

```{r benchmark-eval-2}
bm_sp_sp_all %>%
  getBMRAggrPerformances(as.df = TRUE) %>% 
  arrange(task.id, desc(brier.test.mean))
```

### Visualize

```{r benchmark-eval-3}
plt = plotBMRBoxplots(bm_sp_sp_all, measure = brier, pretty.names = FALSE, 
                      order.lrn = c("classif.ranger.tuned", "classif.svm.tuned", "classif.kknn.tuned",
                                    "classif.gbm.tuned" , "classif.gam.tuned")) +
  #order.lrn = getBMRLearnerIds(bm_sp_sp_all_reduced)) + # gam again duplicated -.-
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") + 
  # scale_color_viridis_d() +
  theme(strip.text.x = element_text(size = 8)) + 
  theme_pubr()

levels(plt$data$task.id) = c("Armillaria", "Diplodia", "Fusarium", "Heterobasidion")
levels(plt$data$learner.id) = c("RF", "SVM", "KNN", "BRT", "GAM")

#plt + ylab("Performance")
plt
```

### Aggregated performances

```{r benchmark-eval-4}
plt2 = plotBMRSummary(bm_sp_sp_all, pretty.names = FALSE) + 
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$task.id) = c("Armillaria", "Diplodia", "Fusarium", "Heterobasidion")
levels(plt2$data$learner.id) = c("RF", "SVM", "KNN", "BRT", "GAM")

plt2
```

### Calculating and visualizing ranks

```{r benchmark-eval-6}
plt3 = plotBMRRanksAsBarChart(bm_sp_sp_all, pos = "tile", 
                              pretty.names = FALSE,
                              order.lrn = c("classif.ranger.tuned", "classif.svm.tuned", "classif.kknn.tuned",
                                            "classif.gbm.tuned" , "classif.gam.tuned")) +
  scale_fill_nejm() +
  theme_pubr()

levels(plt3$data$task.id) = c("Armillaria", "Diplodia", "Fusarium", "Heterobasidion")
levels(plt3$data$learner.id) = c("RF", "SVM", "KNN", "BRT", "GAM")

plt3
```

## Spatial-Non-Spatial

## Non-Spatial-Non-Spatial

## Non-Spatial-No Tuning

## Spatial-No Tuning
