---
title: "Evaluating classifiers on pathogens"
#output: html_notebook
author: "Patrick Schratz, FSU Jena"
---

This report evaluates the performance of the algorithms 

* _Boosted Regression Trees_ (BRT), 
* _Random Forest_ (RF), 
* _Support Vector Machine_ (SVM) 
* _Extreme Gradient Boosting_ (xgboost) 
* _Generalized Linear Model_ (GLM) 
* _Generalized Additive Model_ (GAM) 

on four different pathogens:

- _Armillaria mellea_
- _Diplodia sapinea_
- _Fusarium mellea_
- _Heterobasidion mellea_

```{r setup, include=FALSE}
drake::loadd(bm_sp_sp_all)
# knitr::opts_chunk$set(eval = FALSE)
```

# Print benchmark results

```{r benchmark-eval-2}
getBMRAggrPerformances(bm_sp_sp_all, as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

# Visualize

```{r benchmark-eval-3}
plt = plotBMRBoxplots(bm_sp_sp_all, measure = brier, pretty.names = FALSE, 
  order.lrn = getBMRLearnerIds(bm_sp_sp_all)) +
  # aes(color = learner.id) +
  # scale_color_viridis_d() +
  theme(strip.text.x = element_text(size = 8)) + 
  theme_pubr()

levels(plt$data$task.id) = c("Armillaria", "Diplodia", "Fusarium", "Heterobasidion")
levels(plt$data$learner.id) = c("RF", "SVM", "KNN", "BRT", "GLM", "GAM")

plt + ylab("Performance")
```

## Aggregated performances

```{r benchmark-eval-4}
plt2 = plotBMRSummary(bm_sp_sp_all) + 
  theme_pubr()

levels(plt2$data$task.id) = c("Armillaria", "Diplodia", "Fusarium", "Heterobasidion")
levels(plt$data$learner.id) = c("RF", "SVM", "KNN", "BRT", "GLM", "GAM")

plt2
```

## Calculating and visualizing ranks

```{r benchmark-eval-5}
m = convertBMRToRankMatrix(bm_sp_sp_all, brier)
as_tibble(m)
```

```{r benchmark-eval-6}
plt3 = plotBMRRanksAsBarChart(bm_sp_sp_all, pos = "tile", order.lrn = getBMRLearnerIds(bm_sp_sp_all)) +
  theme_pubr()

levels(plt3$data$task.id) = c("Armillaria", "Diplodia", "Fusarium", "Heterobasidion")
levels(plt$data$learner.id) = c("RF", "SVM", "KNN", "BRT", "GLM", "GAM")

plt3
```

