---
title: "Prediction maps of classifiers"
output: html_document
author: "Patrick Schratz, FSU Jena"
---

Models that did not converge:

- Armillaria: XGBOOST (sp/sp)
- Armillaria: GAM     (sp/sp)

Since we got no performance estimate of these combinations, we did not create prediction maps.

Furthermore, XGBOOST cannot handle new levels factors of the prediction data (or at least I do not know how to solve this).
Therefore, no prediction maps for XGBOOST as the prediction values are all NA.

```{r setup, include=FALSE}
knitr::opts_knit$set(base.dir = 'analysis/rmd/')
knitr::opts_chunk$set(fig.retina = 3, fig.align = 'center', fig.width = 8.5, 
                      fig.height = 5.5, out.width = "100%",
                      fig.path = "../../analysis/paper/submission/3/appendices/")
```

```{r echo = FALSE}
loadd(maps_gam_diplodia, maps_glm_diplodia, 
      maps_gam_fusarium, maps_glm_fusarium,
      #maps_gam_armillaria, 
      maps_glm_armillaria, 
      maps_gam_heterobasidion, maps_glm_heterobasidion, 
      
      maps_svm, maps_rf,
      # maps_xgboost, 
      maps_kknn, 
      maps_brt,
      maps_debugging_diplodia_no_temp)
```

# Pathogens {.tabset .tabset-fade}

## Armillaria

### GAM

```{r prediction-map-armillaria-gam }
NA
```

### GLM

```{r prediction-map-armillaria-glm }
maps_glm_armillaria
```

### BRT

```{r prediction-map-armillaria-brt }
maps_brt[[1]]
```

### RF

```{r prediction-map-armillaria-rf }
maps_rf[[1]]
```

### SVM

```{r prediction-map-armillaria-svm }
maps_svm[[1]]
```

### KKNN

```{r prediction-map-armillaria-knn }
maps_kknn[[1]]
```

### XGBOOST

```{r prediction-map-armillaria-xgboost }
NA
```


## Heterobasidion

### GAM

```{r prediction-map-heterobasidion-gam }
maps_gam_heterobasidion
```

### GLM

```{r prediction-map-heterobasidion-glm }
maps_glm_heterobasidion
```

### BRT

```{r prediction-map-heterobasidion-brt }
maps_brt[[2]]
```

### RF

```{r prediction-map-heterobasidion-rf }
maps_rf[[2]]
```

### SVM

```{r prediction-map-heterobasidion-svm }
maps_svm[[2]]
```

### KKNN

```{r prediction-map-heterobasidion-knn }
maps_kknn[[2]]
```

### XGBOOST

```{r prediction-map-heterobasidion-xgboost }
NA
```


## Diplodia

### GAM

```{r prediction-map-diplodia-gam }
maps_gam_diplodia
```

### GLM

```{r prediction-map-diplodia-glm }
maps_glm_diplodia
```

### BRT

```{r prediction-map-diplodia-brt }
maps_brt[[3]]
```

### RF

```{r prediction-map-diplodia-rf }
maps_rf[[3]]
```

### SVM

```{r prediction-map-diplodia-svm }
maps_svm[[3]]
```

### KKNN

```{r prediction-map-diplodia-knn }
maps_kknn[[3]]
```

### XGBOOST

```{r prediction-map-diplodia-xgboost }
NA
```

### Debugging

```{r prediction-map-diplodia-no-temp }
maps_debugging_diplodia_no_temp
```

## Fusarium

### GAM

```{r prediction-map-fusarium-gam }
maps_gam_fusarium
```

### GLM

```{r prediction-map-fusarium-glm }
maps_glm_fusarium
```

### BRT

```{r prediction-map-fusarium-brt }
maps_brt[[4]]
```

### RF

```{r prediction-map-fusarium-rf }
maps_rf[[4]]
```

### SVM

```{r prediction-map-fusarium-svm }
maps_svm[[4]]
```

### KKNN

```{r prediction-map-fusarium-knn }
maps_kknn[[4]]
```

### XGBOOST

```{r prediction-map-fusarium-xgboost }
NA
```
