---
title: "03_spatial_autocorrelation"
output: html_document
---

# Load packages

```{r}
library(mgcv)
library(gstat)
library(dplyr)
library(MASS)
library(pgirmess)
```

# Load data

```{r}
armillaria_data <- readRDS("/data/marc/mod/survey_data/armillaria_data.rda")
```

# GLM

## Model

```{r}
model_glm <-
  glm(armillaria ~ elevation + slope + soil + ph + lithology + temperature + pisr + precipitation + hail_probability,
    family = binomial(link = "logit"), data = armillaria_data
  )
```

## Variogram

```{r}
glm_resid <-
  armillaria_data %>%
  dplyr::select(x, y) %>%
  cbind(resid = residuals(model_glm)) %>%
  na.omit()
```

```{r}
variogram(resid ~ 1, data = glm_resid, locations = ~ x + y, cutoff = 20000) %>%
  plot()
```

Results: Standardized residuals show spatial autocorrelation within a range of ~ 5000m

# glmmPQL example

## Preperation

Define the same group factor for all observations otherwise glmmPQL does not work.
According to Dormann et al. 2017 Appendix (http://www.ecography.org/sites/ecography.org/files/appendix/e5171.pdf) this produces identical results like official spatial GLMMs.

```{r}
armillaria_data <-
  factor(rep("a", nrow(armillaria_data))) %>%
  as_tibble() %>%
  dplyr::select(group = value) %>%
  bind_cols(armillaria_data)
```

## Model 

glmmPQL model with spatial correlation structure

```{r}
model_glmmPQL <- glmmPQL(armillaria ~ elevation + slope + soil + ph + lithology + temperature + pisr + precipitation + hail_probability,
  random = ~ 1 | group,
  family = binomial(link = "logit"),
  data = armillaria_data,
  correlation = corExp(form = ~ x + y)
)
```


glmmPQL model without spatial correlation structure

```{r}
model_glmmPQL_2 <- glmmPQL(armillaria ~ elevation + slope + soil + ph + lithology + temperature + pisr + precipitation + hail_probability,
  random = ~ 1 | group,
  family = binomial(link = "logit"),
  data = armillaria_data,
)
```

glmmPQL model with spatial correlation structure and manually set range and nugget

```{r}
model_glmmPQL_3 <- glmmPQL(armillaria ~ elevation + slope + soil + ph + lithology + temperature + pisr + precipitation + hail_probability,
  random = ~ 1 | group,
  family = binomial(link = "logit"),
  data = armillaria_data,
  correlation = corExp(10000, form = ~ x + y, nugget = TRUE)
)
```
## Residuals

```{r}
glmmPQL_resid <-
  armillaria_data %>%
  dplyr::select(x, y) %>%
  cbind(resid = residuals(model_glmmPQL, type = "normalized"))
```

```{r}
glmmPQL_2_resid <-
  armillaria_data %>%
  dplyr::select(x, y) %>%
  cbind(resid = residuals(model_glmmPQL_2, type = "normalized"))
```

```{r}
glmmPQL_3_resid <-
  armillaria_data %>%
  dplyr::select(x, y) %>%
  cbind(resid = residuals(model_glmmPQL_3, type = "normalized"))
```

## Variogram

Variogram of the normalized residuals of the glmmPQL model without spatial 
correlation structure. Cut off is set at 20000 m.


```{r}
variogram(resid ~ 1, data = glmmPQL_2_resid, locations = ~ x + y, cutoff = 20000) %>%
  plot(vgm(0.8, "Exp", 10000, nugget = 0.4))
```

Variogram of the normalized residuals of the glmmPQL model with spatial 
correlation structure and automatically fitted variogram parameters. Cut off is set at 20000 m.

```{r}
variogram(resid ~ 1, data = glmmPQL_resid, locations = ~ x + y, cutoff = 20000) %>%
  plot()
```

Variogram of the normalized residuals of the glmmPQL model with spatial 
correlation structure and manually variogram parameters. Cut off is set at 20000 m.

```{r}
variogram(resid ~ 1, data = glmmPQL_3_resid, locations = ~ x + y, cutoff = 20000) %>%
  plot()
```

##Correlogram

Correlogram of the normalized residuals of the glmmPQL model without spatial 
correlation structure.

```{r}
correlog(coords = glmmPQL_2_resid[, 1:2], z = glmmPQL_2_resid[, 3], method = "Moran") %>%
  plot()
```

Correlogram of the normalized residuals of the glmmPQL model with spatial 
correlation structure and manually set variogram parameters.

```{r}
correlog(coords = glmmPQL_3_resid[, 1:2], z = glmmPQL_3_resid[, 3], method = "Moran") %>%
  plot()
```




# gamm example

## Model 1

Convergence error: singular convergence

```{r}
model_gam_1 <- gamm(armillaria ~ s(elevation) + s(slope) + soil + s(ph) + lithology + s(temperature) + s(pisr) + s(precipitation) + (hail_probability),
  family = binomial(link = "logit"), data = armillaria_data
)
```

## Model 2

Convergence error: false convergence

```{r}
model_gam_2 <- gamm(armillaria ~ s(elevation) + s(slope) + soil + s(ph) + lithology + s(temperature) + s(pisr) + s(precipitation) + (hail_probability),
  family = binomial(link = "logit"), data = armillaria_data, correlation = corExp(form = ~ x + y)
)
```


