---
title: "Evaluating classifier performance on pathogens"
output: workflowr::wflow_html
author: "Patrick Schratz, Friedrich-Schiller-University Jena"
bibliography: ../bibliography.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(base.dir = "analysis/rmd/")
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  fig.path = "../../../analysis/paper/submission/3/appendices/",
  echo = FALSE
)

loadd(
  bm_all_pathogens_sp_sp, bm_all_pathogens_sp_nsp,
  bm_all_pathogens_nsp_nsp, bm_all_pathogens_sp_non,
  bm_all_pathogens_nsp_non,

  no_extract_bm_sp_sp_armillaria
)
```

# Introduction

This document shows the predictive performances for the possible infection risk of trees in the Basque Country by the following pathogens:

- _Armillaria mellea_
- _Diplodia sapinea_
- _Fusarium circinatum_
- _Heterobasidion annosum_

The following algorithms were benchmarked:

- Boosted Regression Trees (BRT)
- Generalized Additive Model (GAM)
- Generalized Linear Model (GLM)
- k-Nearest Neighbor (KNN)
- Random Forests (RF)
- Support Vector Machine (SVM)
- Extreme Gradient Boosting (XGBOOST)

The following "algorithm-pathogen-resampling" settings did not return a performance result during performance evaluation.
Due to the missing performance result for these combinations no prediction maps were created.

```{r}
my_tbl <- tibble::tribble(
  ~Pathogen, ~Algorithm, ~Resampling,
  "Armillaria", "XGBOOST", "spatial/spatial",
  "Armillaria", "GAM", "spatial/spatial"
)

kable_styling(
  kable(my_tbl,
    digits = 3, row.names = FALSE, align = "c",
    caption = NULL, format = "html"
  ),
  bootstrap_options = c("striped", "hover", "condensed"),
  position = "center", full_width = FALSE
)
```

Due to two missing results for the pathogen _Armillaria mellea_, this pathogen could not be included in the plots comparing all pathogens against each other.

# Resampling Strategies

The abbreviations of the tabbed resampling strategies follow the scheme:

`<outer resampling> / <inner resampling>`

For example, setting "Spatial-Spatial" means that in both levels a "spatial cross-validation" (@Brenning2012) has been applied.

The inner resampling refers to the hyperparameter tuning level of the nested cross-validation that was applied.

# Results structure

The structure of the following results presentation is as follows:

- Table view of all performances for each resampling setting
- Boxplot comparison for each pathogen and algorithm
- Aggregated performances for each pathogen and algorithm

# Resampling strategies {.tabset .tabset-fade}

## Spatial-Spatial

```{r benchmark-eval-all-2 }
bm_all_pathogens_sp_sp %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-all-3 }
plt <- plotBMRBoxplots(bm_all_pathogens_sp_sp,
  measure = brier, pretty.names = F,
  order.lrn = c(
    "classif.gam.tuned", 
    "classif.kknn.tuned",
    "classif.ranger.tuned",
    "classif.svm.tuned", 
    "classif.gbm.tuned",
    "classif.xgboost.tuned"
  )
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8),
        axis.text.x = element_text(size = 7)
        )

levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt
```

### Aggregated performances

```{r benchmark-eval-all-4 }
plt2 <- plotBMRSummary(bm_all_pathogens_sp_sp, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt2
```

## Spatial-Non-Spatial

```{r benchmark-eval-all-5 }
bm_all_pathogens_sp_nsp %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-all-6 }
plt <- plotBMRBoxplots(bm_all_pathogens_sp_nsp,
  measure = brier, pretty.names = F,
  order.lrn = c(
    "classif.gam.tuned", 
    "classif.kknn.tuned",
    "classif.ranger.tuned",
    "classif.svm.tuned", 
    "classif.gbm.tuned",
    "classif.xgboost.tuned"
  )
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8),
        axis.text.x = element_text(size = 7)
        )

levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt
```

### Aggregated performances

```{r benchmark-eval-all-7 }
plt2 <- plotBMRSummary(bm_all_pathogens_sp_nsp, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt2
```

## Non-Spatial-Non-Spatial

```{r benchmark-eval-all-8 }
bm_all_pathogens_nsp_nsp %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-all-9 }
plt <- plotBMRBoxplots(bm_all_pathogens_nsp_nsp,
  measure = brier, pretty.names = F,
  order.lrn = c(
    "classif.gam.tuned", "classif.kknn.tuned",
    "classif.ranger.tuned",
    "classif.svm.tuned", "classif.gbm.tuned",
    "classif.xgboost.tuned"
  )
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8),
        axis.text.x = element_text(size = 7)
        )

levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt
```

### Aggregated performances

```{r benchmark-eval-all-10 }
plt2 <- plotBMRSummary(bm_all_pathogens_nsp_nsp, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt2
```

## Non-Spatial-No Tuning

```{r benchmark-eval-all-11 }
bm_all_pathogens_nsp_non %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-all-12 }
plt <- plotBMRBoxplots(bm_all_pathogens_nsp_non,
  measure = brier, pretty.names = F,
  order.lrn = c(
    "classif.gam", 
    "classif.binomial",
    "classif.kknn",
    "classif.ranger",
    "classif.svm", 
    "classif.gbm",
    "classif.xgboost"
  )
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8),
        axis.text.x = element_text(size = 7)
        )

levels(plt$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt
```

### Aggregated performances

```{r benchmark-eval-all-13 }
plt2 <- plotBMRSummary(bm_all_pathogens_nsp_non, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt2
```

## Spatial-No Tuning

```{r benchmark-eval-all-14 }
bm_all_pathogens_sp_non %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-all-15 }
plt <- plotBMRBoxplots(bm_all_pathogens_sp_non,
  measure = brier, pretty.names = F,
  order.lrn = c(
    "classif.gam", 
    "classif.binomial",
    "classif.kknn",
    "classif.ranger",
    "classif.svm", 
    "classif.gbm",
    "classif.xgboost"
  )
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") ++
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8),
        axis.text.x = element_text(size = 7)
        )

levels(plt$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt
```

### Aggregated performances

```{r benchmark-eval-all-16 }
plt2 <- plotBMRSummary(bm_all_pathogens_sp_non, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt2
```

# References
