---
title: "Evaluating classifier performance on _Diplodia sapinea_"
output: workflowr::wflow_html
author: "Patrick Schratz, Friedrich-Schiller-University Jena"
bibliography: bibliography.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

source(here::here("code/99-packages.R"))
sourceDirectory("R")

loadd(
  no_extract_bm_sp_sp_diplodia, no_extract_bm_sp_nsp_diplodia,
  no_extract_bm_nsp_nsp_diplodia, no_extract_bm_nsp_non_diplodia,
  no_extract_bm_sp_non_diplodia, diplodia_data
)
```

# Introduction

This document shows the predictive performances for the possible infection risk of trees in the Basque Country by the pathogen _Diplodia sapinea_.

The following algorithms were benchmarked:

- Boosted Regression Trees (BRT)
- Generalized Additive Model (GAM)
- Generalized Linear Model (GLM)
- k-Nearest Neighbor (KNN)
- Random Forests (RF)
- Support Vector Machine (SVM)
- Extreme Gradient Boosting (XGBOOST)

# Resampling Strategies

The abbreviations of the tabbed resampling strategies follow the scheme:

`<outer resampling> / <inner resampling>`

For example, setting "Spatial-Spatial" means that in both levels a "spatial cross-validation" (@Brenning2012) has been applied.

The inner resampling refers to the hyperparameter tuning level of the nested cross-validation that was applied.

# Results structure

Teh results are structured as follows:

- Results table of all performances for each resampling setting
- Boxplot comparison across all algorithms
- Aggregated performances across all algorithms

Even though GBM shows the best performance here, RF was the winning model as long as predictor `pH` was included in the model.
This predictor was removed after causing block artifacts in the prediction maps.
The following results do not include predictor `pH`.

```{r echo = FALSE}
tibble::as_tibble(diplodia_data)
```

# Performance results {.tabset .tabset-fade}

## Spatial-Spatial

```{r benchmark-eval-diplodia-2 }
no_extract_bm_sp_sp_diplodia %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-diplodia-3 }
plt <- plotBMRBoxplots(no_extract_bm_sp_sp_diplodia,
  measure = brier, pretty.names = T,
  order.lrn = getBMRLearnerIds(no_extract_bm_sp_sp_diplodia)
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8), legend.position = "none") 

levels(plt$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT")

plt
```

### Aggregated performances

```{r benchmark-eval-diplodia-4 }
plt2 <- plotBMRSummary(no_extract_bm_sp_sp_diplodia, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$task.id) <- c("Diplodia")
levels(plt2$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT")

plt2
```

## Spatial-Non-Spatial

```{r benchmark-eval-diplodia-5 }
no_extract_bm_sp_nsp_diplodia %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-diplodia-6 }
plt <- plotBMRBoxplots(no_extract_bm_sp_nsp_diplodia,
  measure = brier, pretty.names = T,
  order.lrn = getBMRLearnerIds(no_extract_bm_sp_nsp_diplodia)
) + #
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme(strip.text.x = element_text(size = 8)) +
  theme_pubr()

levels(plt$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT")

plt
```

### Aggregated performances

```{r benchmark-eval-diplodia-7 }
plt2 <- plotBMRSummary(no_extract_bm_sp_nsp_diplodia, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$task.id) <- c("Diplodia")
levels(plt2$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT")

plt2
```

## Non-Spatial-Non-Spatial

```{r benchmark-eval-diplodia-8 }
no_extract_bm_nsp_nsp_diplodia %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-diplodia-9 }
plt <- plotBMRBoxplots(no_extract_bm_nsp_nsp_diplodia,
  measure = brier, pretty.names = T,
  order.lrn = getBMRLearnerIds(no_extract_bm_nsp_nsp_diplodia)
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme(strip.text.x = element_text(size = 8)) +
  theme_pubr()

levels(plt$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT")

plt
```

### Aggregated performances

```{r benchmark-eval-diplodia-10 }
plt2 <- plotBMRSummary(no_extract_bm_nsp_nsp_diplodia, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$task.id) <- c("Diplodia")
levels(plt2$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT")

plt2
```

## Non-Spatial-No Tuning

```{r benchmark-eval-diplodia-11 }
no_extract_bm_nsp_non_diplodia %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-diplodia-12 }
plt <- plotBMRBoxplots(no_extract_bm_nsp_non_diplodia,
  measure = brier, pretty.names = T,
  order.lrn = getBMRLearnerIds(no_extract_bm_nsp_non_diplodia)
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme(strip.text.x = element_text(size = 8)) +
  theme_pubr()

levels(plt$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT")

plt
```

### Aggregated performances

```{r benchmark-eval-diplodia-13 }
plt2 <- plotBMRSummary(no_extract_bm_nsp_non_diplodia, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT")

plt2
```

## Spatial-No Tuning

```{r benchmark-eval-diplodia-14 }
no_extract_bm_sp_non_diplodia %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(brier.test.mean))
```

### Boxplot comparison

```{r benchmark-eval-diplodia-15 }
plt <- plotBMRBoxplots(no_extract_bm_sp_non_diplodia,
  measure = brier, pretty.names = T,
  order.lrn = getBMRLearnerIds(no_extract_bm_sp_non_diplodia)
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme(strip.text.x = element_text(size = 8)) +
  theme_pubr()

levels(plt$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT")

plt
```

### Aggregated performances

```{r benchmark-eval-diplodia-16 }
plt2 <- plotBMRSummary(no_extract_bm_sp_non_diplodia, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

levels(plt2$data$task.id) <- c("Diplodia")
levels(plt$data$learner.id) <- c("GAM", "GLM", "KNN", "RF", "SVM", "BRT")

plt2
```

# Boxplot comparison of all algorithm/tuning settings

```{r aggregate-by-rep }
all_diplodia <- suppressWarnings(map2(
  list(
    no_extract_bm_sp_sp_diplodia,
    no_extract_bm_sp_nsp_diplodia,
    no_extract_bm_nsp_nsp_diplodia,
    no_extract_bm_sp_non_diplodia,
    no_extract_bm_nsp_non_diplodia
  ),
  c(
    "Spatial/Spatial", "Spatial/Non-Spatial", "Non-Spatial/Non-Spatial",
    "Spatial/No Tuning", "Non-Spatial/No Tuning"
  ), ~ {
    getBMRPerformances(.x, as.df = TRUE) %>%
      dplyr::mutate(setting = .y) %>%
      dplyr::mutate(rep = case_when(
        (iter >= 1 & iter <= 5) ~ 1,
        (iter >= 6 & iter <= 10) ~ 2,
        (iter >= 11 & iter <= 15) ~ 3,
        (iter >= 16 & iter <= 20) ~ 4,
        (iter >= 21 & iter <= 25) ~ 5,
        (iter >= 26 & iter <= 30) ~ 6,
        (iter >= 31 & iter <= 35) ~ 7,
        (iter >= 36 & iter <= 40) ~ 8,
        (iter >= 41 & iter <= 45) ~ 9,
        (iter >= 46 & iter <= 50) ~ 10,
        (iter >= 51 & iter <= 55) ~ 11,
        (iter >= 56 & iter <= 60) ~ 12,
        (iter >= 61 & iter <= 65) ~ 13,
        (iter >= 66 & iter <= 70) ~ 14,
        (iter >= 71 & iter <= 75) ~ 15,
        (iter >= 76 & iter <= 80) ~ 16,
        (iter >= 81 & iter <= 85) ~ 17,
        (iter >= 86 & iter <= 90) ~ 18,
        (iter >= 91 & iter <= 95) ~ 19,
        (iter >= 96 & iter <= 100) ~ 20,
        (iter >= 101 & iter <= 105) ~ 21,
        (iter >= 106 & iter <= 110) ~ 22,
        (iter >= 111 & iter <= 115) ~ 23,
        (iter >= 116 & iter <= 120) ~ 24,
        (iter >= 121 & iter <= 125) ~ 25,
        (iter >= 126 & iter <= 130) ~ 26,
        (iter >= 131 & iter <= 135) ~ 27,
        (iter >= 136 & iter <= 140) ~ 28,
        (iter >= 141 & iter <= 145) ~ 29,
        (iter >= 146 & iter <= 150) ~ 30,
        (iter >= 151 & iter <= 155) ~ 31,
        (iter >= 156 & iter <= 160) ~ 32,
        (iter >= 161 & iter <= 165) ~ 33,
        (iter >= 166 & iter <= 170) ~ 34,
        (iter >= 171 & iter <= 175) ~ 35,
        (iter >= 176 & iter <= 180) ~ 36,
        (iter >= 181 & iter <= 185) ~ 37,
        (iter >= 186 & iter <= 190) ~ 38,
        (iter >= 191 & iter <= 195) ~ 39,
        (iter >= 196 & iter <= 200) ~ 40,
        (iter >= 201 & iter <= 205) ~ 41,
        (iter >= 206 & iter <= 210) ~ 42,
        (iter >= 211 & iter <= 215) ~ 43,
        (iter >= 216 & iter <= 220) ~ 44,
        (iter >= 221 & iter <= 225) ~ 45,
        (iter >= 226 & iter <= 230) ~ 46,
        (iter >= 231 & iter <= 235) ~ 47,
        (iter >= 236 & iter <= 240) ~ 48,
        (iter >= 241 & iter <= 245) ~ 49,
        (iter >= 246 & iter <= 250) ~ 50,
        (iter >= 251 & iter <= 255) ~ 51,
        (iter >= 256 & iter <= 260) ~ 52,
        (iter >= 261 & iter <= 265) ~ 53,
        (iter >= 266 & iter <= 270) ~ 54,
        (iter >= 271 & iter <= 275) ~ 55,
        (iter >= 276 & iter <= 280) ~ 56,
        (iter >= 281 & iter <= 285) ~ 57,
        (iter >= 286 & iter <= 290) ~ 58,
        (iter >= 291 & iter <= 295) ~ 59,
        (iter >= 296 & iter <= 300) ~ 60,
        (iter >= 301 & iter <= 305) ~ 61,
        (iter >= 306 & iter <= 310) ~ 62,
        (iter >= 311 & iter <= 315) ~ 63,
        (iter >= 316 & iter <= 320) ~ 64,
        (iter >= 321 & iter <= 325) ~ 65,
        (iter >= 326 & iter <= 330) ~ 66,
        (iter >= 331 & iter <= 335) ~ 67,
        (iter >= 336 & iter <= 340) ~ 68,
        (iter >= 341 & iter <= 345) ~ 69,
        (iter >= 346 & iter <= 350) ~ 70,
        (iter >= 351 & iter <= 355) ~ 71,
        (iter >= 356 & iter <= 360) ~ 72,
        (iter >= 361 & iter <= 365) ~ 73,
        (iter >= 366 & iter <= 370) ~ 74,
        (iter >= 371 & iter <= 375) ~ 75,
        (iter >= 376 & iter <= 380) ~ 76,
        (iter >= 381 & iter <= 385) ~ 77,
        (iter >= 386 & iter <= 390) ~ 78,
        (iter >= 391 & iter <= 395) ~ 79,
        (iter >= 396 & iter <= 400) ~ 80,
        (iter >= 401 & iter <= 405) ~ 81,
        (iter >= 406 & iter <= 410) ~ 82,
        (iter >= 411 & iter <= 415) ~ 83,
        (iter >= 416 & iter <= 420) ~ 84,
        (iter >= 421 & iter <= 425) ~ 85,
        (iter >= 426 & iter <= 430) ~ 86,
        (iter >= 431 & iter <= 435) ~ 87,
        (iter >= 436 & iter <= 440) ~ 88,
        (iter >= 441 & iter <= 445) ~ 89,
        (iter >= 446 & iter <= 450) ~ 90,
        (iter >= 451 & iter <= 455) ~ 91,
        (iter >= 456 & iter <= 460) ~ 92,
        (iter >= 461 & iter <= 465) ~ 93,
        (iter >= 466 & iter <= 470) ~ 94,
        (iter >= 471 & iter <= 475) ~ 95,
        (iter >= 476 & iter <= 480) ~ 96,
        (iter >= 481 & iter <= 485) ~ 97,
        (iter >= 486 & iter <= 490) ~ 98,
        (iter >= 491 & iter <= 495) ~ 99,
        (iter >= 496 & iter <= 500) ~ 100,
        TRUE ~ NA_real_
      )) %>%
      dplyr::group_by(rep, learner.id, setting)
  }
)
)
# now summarise the performance over all groupings

suppressWarnings(
  all_diplodia %<>%
    map_df(~ summarise_at(.x, "brier", mean)) %>%
    ungroup()
)
```

```{r benchmark-eval-diplodia-18 }
all_diplodia %<>%
  mutate(setting = as.factor(setting)) %>%
  mutate(learner.id = fct_recode(learner.id, "GLM" = "classif.binomial")) %>%
  mutate(learner.id = fct_recode(learner.id, "GAM" = "classif.gam.tuned")) %>%
  mutate(learner.id = fct_recode(learner.id, "GAM" = "classif.gam")) %>%
  mutate(learner.id = fct_recode(learner.id, "RF" = "classif.ranger")) %>%
  mutate(learner.id = fct_recode(learner.id, "RF" = "classif.ranger.tuned")) %>%
  mutate(learner.id = fct_recode(learner.id, "SVM" = "classif.svm.tuned")) %>%
  mutate(learner.id = fct_recode(learner.id, "SVM" = "classif.svm")) %>%
  mutate(learner.id = fct_recode(learner.id, "BRT" = "classif.gbm.tuned")) %>%
  mutate(learner.id = fct_recode(learner.id, "BRT" = "classif.gbm")) %>%
  mutate(learner.id = fct_recode(learner.id, "KNN" = "classif.kknn.tuned")) %>%
  mutate(learner.id = fct_recode(learner.id, "KNN" = "classif.kknn"))
```

```{r cv_boxplots_final_brier, fig.width = 6.93, fig.height = 6.13, dev = c("png", "pdf")}
ggplot(all_diplodia, aes(learner.id, brier, fill = setting)) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_fill_nejm() +
  labs(y = "Brier score", x = "") +
  coord_cartesian(ylim = c(0.05, 0.30)) +
  scale_y_continuous(breaks = seq(0.05, 0.30, 0.05)) +
  scale_x_discrete(limits = c("GAM", "GLM", "BRT", "KNN", "RF", "SVM")) +
  theme(legend.title = element_text("CV Type")) +
  theme_pubr() +
  theme(
    axis.text = element_text(size = 14),
    plot.margin = unit(c(0.25, 0.2, 0.2, 0.2), "cm"),
    legend.position = "top",
    legend.text = element_text(size = 14),
    axis.title.y = element_text(margin = margin(
      t = 0, r = 15,
      b = 0, l = 0
    ))
  ) +
  guides(fill = guide_legend(
    title = "", nrow = 3, ncol = 2,
    title.theme = element_text(
      angle = -0,
      size = 12
    )
  ))
```

```{r}
files_in = list.files(here("docs/figure/benchmark-diplodia.Rmd/"), 
                     pattern = "cv_boxplots_final_brier", 
                     full.names = TRUE)
file_copy(files_in,
          rep(here("code/07-paper/submission/3/appendices/"), length(files_in)),
          overwrite = TRUE)
```

```{r echo = FALSE}
file_move(list.files(here("code/07-paper/submission/3/appendices"), 
                     pattern = "cv_boxplots_final_brier", 
                     full.names = TRUE),
          here("code/07-paper/submission/3/latex-source-files/"))
```

# References
