---
title: "Prediction maps of pathogens from various algorithms"
output: workflowr::wflow_html
author: "Patrick Schratz, Friedrich-Schiller-University Jena"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3, fig.align = "center", fig.width = 8.5,
  fig.height = 5.5, out.width = "100%",
  echo = FALSE
)

source(here::here("code/99-packages.R"))
sourceDirectory("R")

loadd(
  maps_debugging_diplodia_no_temp,
  maps_debugging_diplodia_no_precip,
  maps_debugging_diplodia_no_hail,
  maps_debugging_diplodia_no_ph,
  maps_debugging_diplodia_no_soil,
  maps_debugging_diplodia_no_lithology,
  maps_debugging_diplodia_no_slope,
  maps_debugging_diplodia_no_pisr
)
```

# Introduction

This reports is for debugging purposes only.
Its purpose was to identify the predictors responsbile for the artifacts in the prediction maps.

"ph" was found to be the variable causing the artifacts.

Random Forest was used as the algorithm in this case.

# Prediction Maps {.tabset .tabset-fade}

## No Temperature

```{r prediction-map-diplodia-no-temp }
suppressWarnings(print(maps_debugging_diplodia_no_temp[[1]]))
```

## No Precipitation

```{r prediction-map-diplodia-no-precip }
suppressWarnings(print(maps_debugging_diplodia_no_precip[[1]]))
```

##No hail

```{r prediction-map-diplodia-no-hail }
suppressWarnings(print(maps_debugging_diplodia_no_hail[[1]]))
```

## No ph

```{r prediction-map-diplodia-no-ph }
suppressWarnings(print(maps_debugging_diplodia_no_ph[[1]]))
```

## No soil

```{r prediction-map-diplodia-no-soil }
suppressWarnings(print(maps_debugging_diplodia_no_soil[[1]]))
```

## No lithology

```{r prediction-map-diplodia-no-lithology }
suppressWarnings(print(maps_debugging_diplodia_no_lithology[[1]]))
```

## No slope

```{r prediction-map-diplodia-no-slope }
suppressWarnings(print(maps_debugging_diplodia_no_slope[[1]]))
```

## No pisr

```{r prediction-map-diplodia-no-pisr }
suppressWarnings(print(maps_debugging_diplodia_no_pisr[[1]]))
```
